---
title: "Frito Lay Analysis"
author: "Rayon M"
date: "7/30/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## **Project Objectives**

  1.  Classify Attrition
  2.  Identify Trends within the Departments and Job Roles
  3.  Predict Monthly Income

## **Methodology**

  This Analysis was conducted using the following methodology:
  
  1.  Loaded the data set provided (CaseStudy2-data.csv)
  2.  Performed Exploratory Data Analysis to identify key factors that leads to Attrition, and find explanatory variables for the Linear Regression Model.
  3.  Used the Naive Bayes to achieve Classification
  4.  Used Linear Regression to Predict the Monthly Income

  Categorical Data:
  
  *  Attrition
  *  BusinessTravel
  *  Department
  *  EducationField
  *  Gender
  *  JobRole
  *  MaritalStatus
  *  Over18
  *  OverTime
  
  
  Continuous Variables:
  
  *  Age
  *  DailyRate
  *  DistanceFromHome
  *  Education
  *  EmployeeCount
  *  EmployeeNumber
  *  EnvironmentSatisfaction
  *  HourlyRate
  *  JobInvolvement
  *  JobLevel
  *  JobSatisfaction
  *  MonthlyIncome
  *  MonthlyRate
  *  NumCompaniesWorked
  *  PercentSalaryHike
  *  PerformanceRating
  *  RelationshipSatisfaction
  *  StandardHours
  *  StockOptionLevel
  *  TotalWorkingYears
  *  TrainingTimesLastYear
  *  WorkLifeBalance
  *  YearsAtCompany
  *  YearsInCurrentRole
  *  YearsSinceLastPromotion
  *  YearsWithCurrManager


```{r include=FALSE}

# Import the relevant packages and libraries used to perform the analysis.

library(tidyverse)
library(GGally)
library(naniar)
library(dplyr)
library(class)
library(caret)
library(e1071)
library(ggplot2)
library(plotly)
library(data.table)
library(ggthemes)
library(lessR)
library(scales)
library(wesanderson)
```

### *Import Data*
```{r echo = FALSE}
# Read in the employee dataset that will be used in the Analysis.

employee_data = read.csv("C:/Users/Rayon/OneDrive/Documents/Doing DataScience/Doing Data Science/MSDS_6306_Doing-Data-Science/Unit 14 and 15 Case Study 2/CaseStudy2-data.csv")

```

## **Exploratory Data Analysis (EDA)**
###  Explore factors that can lead voluntary employee turnover (Attrition). Identify the top three factors that contribute to turnover.
###  The business is also interested in learning about any job role specific trends that may exist in the data set. You can also provide any other interesting trends and observations from your analysis. 

```{r}
# Look at the structure of the data
str(employee_data) # 870 observations and 36 variables

# display top 6 observations in the data
head(employee_data)


# view the variables in the dataset to determine if there are any "NAs" or missing values.
gg_miss_var(employee_data)

# The dataset did not have any missing  or NA values. 

```

# What trends exist within the data? What is the split between Attrition Yes and No?

```{r}

attrition_split <- data.frame(employee_data %>% group_by(Attrition) %>% summarise(Count = n()))
attrition_split

# The  "Attrition" variable contains 730 No and 140 Yes observations in the dataset.

# Calculate the percentage difference of the No vs. Yes observations.
# Use setDT from the data.table package to calculate the %.

attrition_percent <- setDT (employee_data)[, 100 * .N/ nrow(employee_data), by = Attrition]

attrition_percent$V1 = round(attrition_percent$V1)

ggplot(data = attrition_percent, mapping = aes(x = Attrition, y = V1, fill = Attrition))+
  geom_bar(stat = "identity") +
  geom_text(aes(Attrition, V1-10, label = V1), data = attrition_percent)+
  xlab("Attriton")+
  ylab("Percentages")+
  ggtitle("Attrition Distribution by percent")+
  theme_tufte()+
  theme(legend.position = "none")


# Another option for displaying percentages
BarChart(Attrition, fill = (count) , data = employee_data)

# Summary:
# The data consists of 84% No's and 16% Yes's as it pertains to Attrition.

```

#  Find where Attrition is highest among the Categorical Variables

```{r}

# Attrition vs. Overtime

overtime_percent <- setDT (employee_data)[, 100 * .N/ nrow(employee_data), by = OverTime]

overtime_percent$V1  <-  round(overtime_percent$V1)
colnames(overtime_percent)[2] = "Percent"
overtime_percent <-  data.frame(overtime_percent)
overtime_percent$OverTime <- as.factor(overtime_percent$OverTime)

overtime_percent

# 29% of all employees work Overtime
# 71% of all employees does not work Overtime

ggplotly(ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = OverTime, colour = Attrition))+
  theme_economist()+
   theme(legend.position = "none")+
  ggtitle("Attrition vs. OverTime"))

# 57% of persons who left the company were working overtime! This will be one of the factors that leads to Attrition
# Attrition was highest in the group of employees that worked overtime

# ggplot(data = employee_data, mapping = aes(y = OverTime))+
#   geom_bar()+
#   geom_text(aes(OverTime, label = Percent, Percent = Percent + 1000000), data = overtime_percent)+
#   coord_flip()+
#   scale_y_discrete(expand = c(0, 0))+
#   scale_x_discrete(expand = c(0, 0))


BarChart(OverTime, by = Attrition , data = employee_data)

# Heat map showing that a large number of the employees don't work overtime.

pal <- wes_palette("Zissou1", 100, type = "continuous")
ggplotly(employee_data %>% 
  count(Attrition, OverTime)%>%
  ggplot(mapping = aes(x = Attrition,  y = OverTime)) +
  geom_tile(mapping = aes(fill = n))+
  scale_fill_gradientn(colours = pal) + 
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) + 
  coord_equal()+
  theme_economist()+
  ggtitle("Attrition vs. OverTime"))


mosaicplot(Attrition~OverTime, data = employee_data, col=c("darkslategray4", "orange"), main ="Attrition vs Overtime", xlab = "Attrition", ylab = "Overtime", border = "black", las=0.5, shade=TRUE)


```


```{r}
# Attrition vs BusinessTravel
attrition_businesstravel <- ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = BusinessTravel, colour = Attrition))+
  theme_economist()+
  ggtitle("Attrition vs. Business Travel")
  ggplotly(attrition_businesstravel)

employee_data %>% count(Attrition, BusinessTravel)

# 15% of employees that traveled Rarely left the company
# 12% of employees that were Non-Travel left the company
# 22% of employees that were Travel_Frequently left the company

attrition_businesstravel2 <-  employee_data %>% 
  count(Attrition, BusinessTravel)%>%
  ggplot(mapping = aes(x = Attrition,  y = BusinessTravel)) +
  geom_tile(mapping = aes(fill = n))
  ggplotly(attrition_businesstravel2)
  
# The Highest Attrition percentage was from those who Traveled rarely - 22%. Hence, I don't think business Travel is a factor for Attrition.
  #Business Travel has a lower Median for Attrition than non Attrition
#one department is seeing a high turnover 
```  
  
  
```{r}
# Attrition vs Department

attrition_department <- ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = Department, colour = Attrition))+
  theme_economist()+
  ggtitle("Attrition vs. Department")
  
  ggplotly(attrition_department)

employee_data %>% count(Attrition, Department)

# 17% of those that left were from the Human Resources Department
# 13% of those that left were from the Research & Development Department
# 21% of that that left were from the Sales Department

attrition_department2 <-  employee_data %>% 
  count(Attrition, Department)%>%
  ggplot(mapping = aes(x = Attrition,  y = Department)) +
  geom_tile(mapping = aes(fill = n))+
  theme_economist()+
  ggtitle("Attrition vs. Department")

  ggplotly(attrition_department2)
  
# Do sales people work more overtime?
  
  ggplotly(ggplot(data = employee_data)+
    geom_bar(mapping = aes(x = Department, fill = OverTime, alpha = JobRole), position = "dodge") +
    ggtitle("Overtime by Department")+
    xlab("Department")+
    ylab("Number of Employees")+
    theme_economist()+
      theme(legend.position = "none")
  )

# No. There is a higher percentage of people working overtime in the Research and Development Department - 29%
  #65% of Overtime comes from the Research & Development Department
  #32% of Overtime comes from the Sales Department

PieChart(Department, data = employee_data, hole = 0)

# The Research and Development Department makes up 65% of the company. Department is not one of the Factors that leads to Attrition.

# What is the percentage of non-exempt people leaving the company versus full time?

```


```{r}
# Attrition v. EducationField

ggplotly(ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = EducationField, colour = Attrition))+
  theme_economist()+
  ggtitle("Attrition vs. EducationField"))

employee_data %>% count(Attrition, EducationField)  


```


```{r}
# Attrition v. Gender

ggplotly(ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = Gender, colour = Attrition))+
  theme_economist()+
  ggtitle("Attrition vs. Gender"))

employee_data %>% count(Attrition, Gender)

# There are more males working in the company than females.
# The attrition rate among the two Genders were approximately the same at 15%
# No indicative of Attrition

```


```{r}
# Attrition v. JobRole

ggplotly(ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = JobRole, color = OverTime))+
  theme_economist()+
  ggtitle("Attrition vs. JobRole"))

# JobRole is not a factor for Attrition.

BarChart(JobRole, by=Department, data = employee_data)

employee_data %>% ggplot(mapping = aes(x = JobRole, fill = Attrition))+
  geom_bar()

Attrititon_table <- data.frame(employee_data %>% count(Attrition, JobRole))
Attrititon_table

ggplot(data = Attrititon_table, mapping = aes(x = JobRole, y = n))+
  geom_boxplot()

ggplot(data = employee_data, aes(x = JobRole, y= Attrition, color = Attrition))+
  geom_count()+
  coord_flip()


```


```{r}
# Attrition v. MaritalStatus

ggplotly(ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = MaritalStatus, color = Attrition))+
  theme_economist()+
  ggtitle("Attrition vs. MaritalStatus"))

employee_data %>% count(Attrition, MaritalStatus)

# Higher Attrition count among Single people

70+199
70/269*100 # 26%

70/140*100 #50%

58+352
58/410*100 # 14%

58/140*100 # 41%

12+179
12/179*100 # 6%, 

12/140*100 # 8 %


ggplotly(ggplot(data = employee_data, mapping =  aes(x = OverTime, y = MaritalStatus, fill = Attrition))+
  geom_count()+
  coord_flip()
)

employee_maritalstatus <- employee_data[,c(3,19)]

view(employee_data)

mosaicplot(Attrition~MaritalStatus, data = employee_data, col=c("Blue", "red4", "Green"), main ="Attrition vs MaritalStatus", xlab = "Attrition", ylab = "MaritalStatus", las =1, border = "chocolate", shade = TRUE)



```


```{r}
# Attrition v. Over18
ggplotly(ggplot(data = employee_data) + 
  geom_count(mapping = aes(x = Attrition, y = Over18, colour = Attrition))+
  theme_economist()+
  ggtitle("Attrition vs. Over18"))


# No significance to the Study. All observations are over18.
```
  
  
# Find where Attrition is highest among the Continuous variables


```{r}

employee_data %>% 
  select (Attrition, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager) %>%
  ggpairs(mapping = aes(color = Attrition))

# Attrition looks to have the lowest median in the "Years in current role" variable and No Attrition have the highest, why?
# Attrition looks to have the lowest median in the "Years with Current Manager" and No Attrition have the highest, why?
# Attrition has a high correlation with "Years In current role" and Years with current manager? why? Overall the Correlation is 0.709 (Strong)
# Attrition has a medium correlation with "Years In current role" and Years since Last Promotion? why?

ggplot(data = employee_data, mapping = aes(x = YearsInCurrentRole, ..density..)) + 
  geom_freqpoly(mapping = aes(color = Attrition),binwidth = 1)

# The lower the "Years In Current Role" variable the more likely there is attrition -- which makes sense because people left before spending a long time. What else is driving them to leave?

ggplot(data = employee_data, mapping = aes(x = YearsInCurrentRole, y = YearsAtCompany)) +
  geom_point(position = "jitter")

```


```{r}
employee_data %>% 
  select(Attrition, YearsAtCompany, YearsInCurrentRole) %>% 
  ggpairs(mapping = aes(color = Attrition))

```

```{r}
employee_data %>% 
  select (Attrition,TrainingTimesLastYear, WorkLifeBalance, YearsAtCompany, YearsInCurrentRole) %>%
  ggpairs(mapping = aes(color = Attrition))

# Years At Company and Years In current Role has a high correlation of 0.776; however, that's normal, as you need to be in the company to change roles. And the longer you stay the more likely.


```


```{r}
employee_data %>% 
  select (Attrition, RelationshipSatisfaction, StandardHours, StockOptionLevel, TotalWorkingYears) %>% 
  ggpairs(mapping = aes(color = Attrition))

# The average total working years is lower for the employees that left than the average working years of employees that stayed. However, the highest value of working years is in the attrition field. Therefore, this is not a candidate for selecting attrition.
# RelationshipSatisfaction is not a factor for attrition. The average for relationship satisfaction is the same for those that stayed versus those that leave. However, 25% of the lower scores is in the Attrition column
#Stock Option is not a factor
#Standard Hours is not a factor


```


```{r}
employee_data %>%
 select(Attrition, PercentSalaryHike,PerformanceRating) %>%
  ggpairs(mapping = aes(color = Attrition))

#High correlation between Percent Salary Hike and Performance Rating. Interesting. Employees only provided a rating of 3 or 4. Why?

employee_data %>% select( Attrition,
  JobSatisfaction
  ,Age
 ,MonthlyIncome           
 ,MonthlyRate
 ,NumCompaniesWorked) %>%
  ggpairs(mapping = aes(color = Attrition))

# NumCompaniesWorked is not a factor for Attrition
# Employees with lower monthly incomes are more like to leave to the company. The average MonthlyIncome was way lower than those who stayed with the company. MonthlyIncome is a factor for Attrition
# MonthlyRate is not a factor for Attrition
# Age is not a factor for Attrition. There were more people within the same age groups that stayed with the company.
# Histogram of Age vs Attriton will prove this.
# Not enough evidence to suggest that JobSatisfaction is a factor for attrition.

employee_data %>% select(
  Attrition
,HourlyRate
,JobInvolvement          
,JobLevel                
,JobRole) %>%
  ggpairs(mapping = aes(color = Attrition))

#Hourly rate is not a factor for Attrition. On average it is higher for those that left than those that stayed

#The median job involvement for males were lower than females

#JobInvolvement
ggplotly(ggplot(data = employee_data, mapping = aes(x = Attrition, y = JobInvolvement, colour = Attrition))+
           geom_boxplot())

employee_data %>% count(JobInvolvement)
BarChart(JobInvolvement, fill = (count) , data = employee_data)
ggplotly(employee_data %>% ggplot(mapping = aes(x = JobInvolvement, fill = Attrition))+
  geom_bar())

# Job involvement is not a factor for Attrition. 
```


```{r}
# JobLevel is a factor for Attrition. As Job level increases, Attrition decreases. More Attrition occurred with folks at the lowest job level.

BarChart(JobLevel, fill = (count) , data = employee_data)
ggplotly(employee_data %>% ggplot(mapping = aes(x = JobLevel, fill = Attrition))+
  geom_bar())

```


```{r}

employee_data %>% select(
  Attrition
 ,Education         
 ,EmployeeCount           
 ,EmployeeNumber          
 ,EnvironmentSatisfaction 
) %>%
  ggpairs(mapping = aes(color = Attrition))

# Education is not a factor for Attrition
# EmployeeCount is not a factor for Attrition
  ggplot(data = employee_data, mapping = aes( x = EmployeeCount))+
  geom_boxplot()

# EmployeeNumber is not a factor for Attrition
  ggplot(data = employee_data, mapping = aes( x = EmployeeNumber))+
  geom_boxplot()

# EmployeeSatisfaction is not a factor for Attrition

```

```{r}

employee_data %>% select(
 Attrition               
 ,DailyRate               
 ,DistanceFromHome
 ,PercentSalaryHike
 ,PerformanceRating
) %>%
  ggpairs(mapping = aes(color = Attrition))

 # DailyRate is not a factor for Attrition             
 # DistanceFromHome is not a factor for Attrition
 # PercentSalaryHike is not a factor for Attrition
 # PerformanceRating is not a factor for Attrition

```

```{r}
ggplot(employee_data, mapping = aes(x=BusinessTravel)) +
  geom_bar(mapping =aes(color = Attrition), position = "dodge")

#Non-Travel, Travel_Frequently, Travel_rarely

ggplot(employee_data, mapping = aes(x=BusinessTravel, y = DistanceFromHome)) +
  geom_boxplot(mapping =aes(color = Attrition))
# people who work on average a further distance from home and don't travel have a high attrition!


ggplot(employee_data, mapping = aes(x=BusinessTravel, y = Department)) +
  geom_tile(mapping =aes())+
  coord_flip()

employee_data %>% 
  count(BusinessTravel, Department) %>%
  ggplot(mapping = aes(x = BusinessTravel, y = Department)) +
  geom_tile(mapping = aes(fill = n))

```

```{r}

employee_data %>% select(Attrition
  ,TrainingTimesLastYear
  ,WorkLifeBalance
  ,YearsAtCompany
  ,YearsInCurrentRole
  ,YearsSinceLastPromotion
  ,YearsWithCurrManager) %>%
  ggpairs(mapping = aes(color = Attrition))


```



# **Classification Model**
## Naive Bayes

```{r}

# Find the average accuracy, sensitivity, and specificity of the model. The sensitivity and specificity must both be above 60%.

master_acc <-  numeric(100)
master_sensitivity  <-  numeric(100)
master_specificivity <-  numeric(100)
master_seed <-  numeric(100)


for(seed in 1:100)
{
  set.seed(seed)
  splitPerc = 0.50
  
  Index = sample(1:dim(employee_data)[1],round(splitPerc * dim(employee_data)[1]))
  Train = employee_data[Index,]
  Test = employee_data[-Index,]
  
  model_attrition_yesorno = naiveBayes(Train[,c("MonthlyIncome", "JobLevel", "OverTime")], factor(Train$Attrition, labels=c("No", "Yes")))
  CM = confusionMatrix(table(factor(Test$Attrition, labels = c("No", "Yes")), predict(model_attrition_yesorno,Test[,c("MonthlyIncome", "JobLevel", "OverTime")])))
  
  master_acc[seed] <-  CM$overall[1]
  master_sensitivity[seed]  <-  CM$byClass[1]
  master_specificivity[seed]  <-  CM$byClass[2]
  master_seed[seed] <-  seed
}

mean(master_seed)
mean(master_acc)
mean(master_sensitivity)
na.omit(max(master_specificivity))

 # model = naiveBayes(train[,c(1,2)],train$Species)
 #  table(predict(model,test[,c(1,2)]),test$Species)
 #  CM = confusionMatrix(table(predict(model,test[,c(1,2)]),test$Species))
 #  masterAcc[j] = CM$overall[1]

# The mean Accuracy is 0.8544061
# The mean Sensitivity is 0.8688525
# The mean Specificity is 0.6470588

```


# Naive Bayes Model
# Final Model used for the Naive Bayes classification

```{r echo=FALSE}

set.seed(5)

splitPerc = 0.70
TrainIndices = sample(1:dim(employee_data)[1],round(splitPerc * dim(employee_data)[1]))
Train = employee_data[TrainIndices,]
Test = employee_data[-TrainIndices,]

model_attrition_yesorno = naiveBayes(Train[,c("MonthlyIncome", "JobLevel", "OverTime")],factor(Train$Attrition, labels=c("No", "Yes")))
CM = confusionMatrix(table(factor(Test$Attrition, labels = c("No", "Yes")), predict(model_attrition_yesorno,Test[,c("MonthlyIncome", "JobLevel", "OverTime")])))

CM

```


## Competition dataset with No Attrition. Use the Naive Bayes Model to predict Attrition.

```{r echo = FALSE}

#Import the competition dataset
comp_dataset <- read.csv("C:/Users/Rayon/OneDrive/Documents/Doing DataScience/Doing Data Science/MSDS_6306_Doing-Data-Science/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Attrition.csv")

# Use the NaiveBayes Classification Model to classify each observation as Yes or No to Attrition.
# Model name: model_attrition_yesorno
# Create a new variable, comp_dataset$PredAttrition to store the predicted values.

comp_dataset$PredAttrition <- predict(model_attrition_yesorno,comp_dataset[,c("MonthlyIncome", "JobLevel", "OverTime")])

# You must provide the labels (ordered by ID) in a csv file.

write.csv(comp_dataset[,c(1,36)], "Case2PredictionsMorris Attrition.csv")

```


## Exploratory Data Analysis for Regression Model

```{r}
# Plot the following variables to see the correlation with MonthlyIncome

employee_data %>% select(MonthlyIncome, Age, Department, EducationField, JobLevel, JobRole, TotalWorkingYears, YearsAtCompany) %>%
ggpairs()

```


## Regression Model to predict Monthly Income
```{r}

# Goal: create a regression model that has an RMSE less than $3000

# Dependent Variable: MonthlyIncome
# Predictors/Experiment variable(s): Age and Total Working Years


numMSPEs = 1000

MSPEHolderModel4 = numeric(numMSPEs)
MSPEHolderModel5 = numeric(numMSPEs)
MSPEHolderModel10 = numeric(numMSPEs)

for (i in 1:numMSPEs)
{
  
  index <- createDataPartition(employee_data$ID, p = .60, list = FALSE)
  monthly_income_train <- employee_data[index, ]
  monthly_income_test <- employee_data[-index, ]
  
  
  #Model 4
  monthly_income_model_fit4 = lm(MonthlyIncome ~ JobLevel+TotalWorkingYears, data = monthly_income_train)
  monthly_income_Pred_4 = predict(monthly_income_model_fit4, newdata = monthly_income_test)
  MSPE4 = mean((employee_data$MonthlyIncome - monthly_income_Pred_4)^2)
  MSPE4
  MSPEHolderModel4[i] = MSPE4
  
  #Model 5
  monthly_income_model_fit5 = lm(MonthlyIncome ~ JobRole, data = monthly_income_train)
  monthly_income_Pred_5 = predict(monthly_income_model_fit5, newdata = monthly_income_test)
  MSPE5 = mean((employee_data$MonthlyIncome - monthly_income_Pred_5)^2)
  MSPE5
  MSPEHolderModel5[i] = MSPE5
  
}

mean(MSPEHolderModel4)
summary(monthly_income_model_fit4)
confint(monthly_income_model_fit4)

mean(MSPEHolderModel5)
summary(monthly_income_model_fit5)
confint(monthly_income_model_fit5)

testMSE4 = mean((monthly_income_test$MonthlyIncome - monthly_income_Pred_4) ^ 2)
testMSE4
RMSE4 <- sqrt(testMSE4)
RMSE4

testMSE5 = mean((monthly_income_test$MonthlyIncome - monthly_income_Pred_5) ^ 2)
testMSE5
RMSE5 <- sqrt(testMSE5)
RMSE5

```

## Competition dataset with No MonthlyIncome. Use the Linear Regression Model to predict Monthly Income.

```{r}
#Load in the competition dataset

monthly_income_comp_dataset = read.csv("C:/Users/Rayon/OneDrive/Documents/Doing DataScience/Doing Data Science/MSDS_6306_Doing-Data-Science/Unit 14 and 15 Case Study 2/CaseStudy2CompSet No Salary.csv")

#Use the Linear Regression Model to predict the MonthlyIncome for each observation.
# Model name = "model_attrition_yesorno"

monthly_income_comp_dataset$PredMonthlyIncome = predict(monthly_income_model_fit4, monthly_income_comp_dataset)


write.csv(monthly_income_comp_dataset[,c(1,36)], "Case2PredictionsMorris Salary.csv")

```



```{r}
employee_data %>% ggplot(mapping = aes(x = MonthlyIncome, y = Department))+
  geom_point()+
  geom_smooth(method = "lm" , se = FALSE)+
  ggtitle("Total Years vs. MonthlyIncome")
#cor.test(employee_data$MonthlyIncome, as.factor(employee_data$Department))


employee_data %>% ggplot(mapping = aes(x = MonthlyIncome, y = Age))+
  geom_point()+
  geom_smooth(method = "lm" , se = FALSE)+
  ggtitle("Total Years vs. MonthlyIncome")


employee_data %>% ggplot(mapping = aes(x = MonthlyIncome, y = EducationField))+
  geom_point()+
  geom_smooth(method = "lm" , se = FALSE)+
  ggtitle("Total Years vs. MonthlyIncome")+
  coord_flip()

employee_data %>% ggplot(mapping = aes(x = MonthlyIncome, y = JobLevel))+
  geom_point()+
  geom_smooth(method = "lm" , se = FALSE)+
  ggtitle("Total Years vs. MonthlyIncome")
cor.test(employee_data$MonthlyIncome, employee_data$JobLevel)


employee_data %>% ggplot(mapping = aes(x = MonthlyIncome, y = JobRole))+
  geom_point()+
  geom_smooth(method = "lm" , se = FALSE)+
  ggtitle("Total Years vs. MonthlyIncome")+
  coord_flip()

str(employee_data)

cor.test(employee_data$MonthlyIncome, employee_data$YearsAtCompany)

employee_data %>% ggplot(mapping = aes(x = TotalWorkingYears, y = YearsAtCompany, position = "jitters"), position = "jitters")+
  geom_point()+
  geom_smooth(method = "lm" , se = FALSE)+
  ggtitle("Total Years vs. MonthlyIncome")


cor.test(employee_data$MonthlyIncome, employee_data$TotalWorkingYears)

employee_data %>% ggplot(mapping = aes(x = TotalWorkingYears, y = MonthlyIncome))+
  geom_point()+
  geom_smooth(method = "lm" , se = FALSE)+
  ggtitle("Total Years vs. MonthlyIncome")+
  coord_flip()



```



#EDA CHUNK
```{r}


ggplot(data = employee_data, mapping = aes(x = MonthlyIncome, colour = JobRole)) +
  geom_freqpoly()

# Observation, the people with salary less than $5,000 is greatest among the Research and Development Department, why?
# Are they younger?, Do they have lower job levels, do they work less hours? Did more people leave the company out of the Research & Development department? What department had the highest Attrition?


# Research Scientist are the highest among those making less than $5,000. They start out a little higher than other roles. but only less than 5 are making over $10,000 
ggplot(data = employee_data, mapping = aes(x = MonthlyIncome, colour = JobRole)) +
  geom_freqpoly()
 
employee_data %>% group_by(JobRole) %>% count(MonthlyIncome)

# Monthly Income
ggplot(data = employee_data, mapping = aes(x= JobRole, y = MonthlyIncome, color = Attrition))+
  geom_boxplot()

#How does overtime affect this? 
ggplot(data = employee_data, mapping = aes(x= JobRole, y = MonthlyIncome, color = OverTime))+
  geom_boxplot()

#How many people get overtime by jobrole? Only 65 research scientist gets overtime.
employee_data %>% group_by(JobRole) %>% count(OverTime)


# What does this look like for the daily rate?
ggplot(data = employee_data, mapping = aes(x= JobRole, y = DailyRate, color = Attrition))+
  geom_boxplot()

# What does this look like for the Monthly rate?
ggplot(data = employee_data, mapping = aes(x= JobRole, y = MonthlyRate, color = Attrition))+
  geom_boxplot()


# Attrition is high among those who make less that 5,000. The graph below shows this.
ggplot(data = employee_data, mapping = aes(x = MonthlyIncome, colour = Attrition)) +
  geom_freqpoly()

# The average monthlyIncome was significantly lower in one Education Field than the others? Why?
# The average monthlyIncome was significantly lower in one Department than the others? Why?


ggplot(data = )
ggplot(data = employee_data, mapping = aes(x= EducationField, y = MonthlyRate, color = Attrition))+
  geom_boxplot()

```


